Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-17 arxiv/proof_same_cov.tex
polish

	\paragraph{Proof of equation \eqref{eq_gA}.}
	To facilitate the analysis, we introduce the following notations.
	Let
		\[ \cE \define {\sum_{k=1}^t \varepsilon^{(k)} A_k^{\top}}, \text{ and } \cW \define U_X U_X^{\top} \cE (AA^{\top})^{-1}. \]
	For any $j = 1,2,\dots, t$, let
	\begin{align*}
		H_j &\define B^{\star} A^{\top} (AA^{\top})^{-1} A_j - \beta^{(j)}, \\
		E_j &\define \cW A_j - \varepsilon^{(j)}.
	\end{align*}
	Then, we have that
	\[ g(A) = \sum_{j=1}^t \bignorm{H_j + E_j}^2. \]
	To compute the expectation of $g(A)$, we first observe that for any first order terms of $\varepsilon^{(1)}, \dots, \varepsilon^{(t)}$, their expectation is zero. Therefore,
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t }{g(A)} = \sum_{j=1}^t \norm{H_j}^2 + \exarg{\set{\varepsilon^{(j)}}_{j=1}^t, X}{\sum_{j=1}^t \norm{E_j}^2}.
	\end{align*}
	Hence, it suffices to show that the second part of the equation above is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	Conditional on $X$, we have that
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\norm{E_j}^2}
		= \exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
	\end{align*}
	We consider both terms in the above equation one by one.
	First, note that
	\begin{align}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE (AA^{\top})^{-1} \cE^{\top}}. \label{eq_proof_same_cov_1}
	\end{align}
	We observe that
	\begin{align*}
			\exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\cE (AA^{\top})^{-1} \cE^{\top}}
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above, the first equation uses the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The second equation uses the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{-1} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
	\begin{align*}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
	\end{align*}
	because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	Next, note that
	\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW \cE^{\top}}, \]
	which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$. Hence the proof is complete.

>>>>

	\paragraph{Proof of equation \eqref{eq_gA}.}
	To facilitate the analysis, we consider the following matrix notations.
	Let
		\[ \cE A^{\top} = {\sum_{k=1}^t \varepsilon^{(k)} A_k^{\top}}, \text{ and } \cW \define X(X^{\top} X)^{\top} X^{\top} \cE A^{\top} (AA^{\top})^{-1}. \]
	For any $j = 1,2,\dots, t$, let
	\begin{align*}
		H_j &\define B^{\star} A^{\top} (AA^{\top})^{-1} A_j - \beta^{(j)}, \\
		E_j &\define \cW A_j - \varepsilon^{(j)}.
	\end{align*}
	Then, we can write the function $g(A)$ conveniently as
	\[ g(A) = \sum_{j=1}^t \bignorm{X H_j + E_j}^2. \]
	To compute the expectation of $g(A)$, we first observe that for any first order terms of $\varepsilon^{(1)}, \dots, \varepsilon^{(t)}$, their expectation is zero. Therefore,
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t, X}{g(A)} = n \sum_{j=1}^t \norm{\Sigma^{1/2} H_j}^2 + \exarg{\set{\varepsilon^{(j)}}_{j=1}^t, X}{\sum_{j=1}^t \norm{E_j}^2}.
	\end{align*}
	Hence, it suffices to show that the second part of the equation above is equal to $\sigma^2 (n\cdot t - p\cdot r)$.
	Conditional on $X$, we have that
	\begin{align*}
		\exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\norm{E_j}^2}
		= \exarg{\set{\varepsilon^{(j)}}_{j=1}^t}{\sum_{j=1}^t \bigbrace{\bignorm{\cW A_j}^2 - 2\inner{\cW A_j}{\varepsilon^{(j)}} }} + \sigma^2 \cdot n \cdot t.
	\end{align*}
	We consider both terms in the above equation one by one.
	First, note that
	\begin{align}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sum_{j=1}^t \bigtr{\cW A_j A_j^{\top} \cW^{\top}} = \bigtr{\cW AA^{\top} \cW^{\top}} = \bigtr{U_X U_X^{\top} \cE A^{\top} (AA^{\top})^{-1} \cE^{\top} A}. \label{eq_proof_same_cov_1}
	\end{align}
	We observe that
	\begin{align*}
			\exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\cE (AA^{\top})^{-1} \cE^{\top}}
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \sum_{k=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_k {\varepsilon^{(k)}}^{\top}} \\
		&= \exarg{\set{\varepsilon^{(j)}}_{i=1}^t}{\sum_{j=1}^t \varepsilon^{(j)} A_j^{\top} (AA^{\top})^{-1} A_j {\varepsilon^{(j)}}^{\top}} \\
		&= \sigma^2 \cdot r \cdot \id_{n\times n}.
	\end{align*}
	In the above, the first equation uses the fact that for any $j\neq k$, $\varepsilon^{(j)}$ and $\varepsilon^{(k)}$ are pairwise independent.
	The second equation uses the fact that $\sum_{j=1}^t A_j^{\top} (AA^{\top})^{-1} A_j = \bigtr{\id_{r\times r}} = r$, and $\ex{\varepsilon^{(j)} {\varepsilon^{(j)}}^{\top}} = \sigma^2 \cdot \id_{n\times n}$.
	Therefore, we have that
	\begin{align*}
		\sum_{j=1}^t \norm{\cW A_j}^2 = \sigma^2 \cdot r \cdot \bigtr{U_X U_X^{\top}} = \sigma^2 \cdot r \cdot p,
	\end{align*}
	because $U_X$ has rank $p$ by Fact \ref{lem_minv}.
	Next, note that
	\[ \sum_{j=1}^t \inner{\cW A_j}{\varepsilon^{(j)}} = \bigtr{\cW \cE^{\top} A}, \]
	which is equal to equation \eqref{eq_proof_same_cov_1} by the definition of $\cW$. Hence the proof is complete.
